Calibration

### 2020-12-23 Paper.Rmd
Code and data extractions to start doing test of the model against real data (landtrendr)

1. The context: MRV and predicting Terrestrial C
2. Status quo: LULC * look up table
  + very coarse, doesn't capture spatiotemporal varibility
  + rightly captures that disturbance and land use change both critical to C accounting and, more importantly, a place where we can exert some control
3. Alternative: SDA
  + models capture process
  + data constrains (data driven, model as cov)
  + data fusion
  + seamless transition from reanalysis to nowcast to forecast [@Dietze2017a]
  + Disturbance is exactly why you would want to nudge a model back to reality
4. The problem: Conventional SDA doesn't capture jumps
  + series of nudges
  + potentially misattributes pools and fluxes
  + "drags down" neighbors **no sims of this yet**
5. The solution:
  + assimilate P(dist)
  + update C | dist
6. This paper
  + Illustration of problem
  + Derivation Multinomial
  + Simulated data experiments (known disturbance)
  + Real-world example

>>>>

1. The context: MRV and predicting Terrestrial C
2. Status quo: LULC * look up table
  + very coarse, doesn't capture spatiotemporal varibility
  + rightly captures that disturbance and land use change both critical to C accounting and, more importantly, a place where we can exert some control
  + BULC
  + state-and-transition modeling
  + process-model projections
3. Alternative: SDA
  + models capture process
  + data constrains (data driven, model as cov)
  + data fusion
  + seamless transition from reanalysis to nowcast to forecast [@Dietze2017a]
  + Disturbance is exactly why you would want to nudge a model back to reality
4. The problem: Conventional SDA doesn't capture jumps
  + series of nudges
  + potentially misattributes pools and fluxes
  + "drags down" neighbors **no sims of this yet**
5. The solution:
  + assimilate P(dist)
  + update C | dist
6. This paper
  + Illustration of problem
  + Derivation Multinomial
  + Simulated data experiments (known disturbance)
  + Real-world example

### 2020-12-23 StateTransition.Rmd
Code and data extractions to start doing test of the model against real data (landtrendr)

  } ## end loop over disturbance class c
} ## end loop over time
TRANS = t(t(tmat)/colSums(tmat))
BBAR  = bioBar/tmat
PBBAR = pbioBar/tmat
BVAR  = sqrt((bioSq -  (bioBar  *  bioBar)/tmat) / (tmat - 1))
PBVAR = sqrt((pbioSq - (pbioBar * pbioBar)/tmat) / (tmat - 1))
BVAR/BBAR - PBVAR
## in all cases, disturbance CV was higher on an absolute than a relative basis
save(PBBAR,PBVAR,file="STprior.RData")
```

@@

>>>>

  } ## end loop over disturbance class c
} ## end loop over time
TRANS = t(t(tmat)/colSums(tmat))
BBAR  = bioBar/tmat
PBBAR = pbioBar/tmat
BVAR  = sqrt((bioSq -  (bioBar  *  bioBar)/tmat) / (tmat - 1))
PBVAR = sqrt((pbioSq - (pbioBar * pbioBar)/tmat) / (tmat - 1))
BVAR/BBAR - PBVAR
## in all cases, disturbance CV was higher on an absolute than a relative basis
colnames(TRANS)<-c("UN","OTHER","CUT","FIRE","PEST")
names(d.prior)[1:5] <- colnames(TRANS)
rownames(TRANS)<-c("water","devel","barren","forest","shrub","---","herb","crop","wetland","NA")
TRANS
save(d.prior,TRANS,PBBAR,PBVAR,file="STprior.RData")
```

@@

coord = xyFromCell(bioS,fsamp[,1])
points = spTransform(SpatialPoints(coord,crs(bioS)),crs("+proj=longlat +datum=WGS84"))

save(points,bextract,cextract,dextract,file="extract.h03v04.RData")
```

### 2020-12-23 methods.Rmd
Code and data extractions to start doing test of the model against real data (landtrendr)

1. Illustration of problem
2. Derivation Multinomial

@@

*Ensemble adjustment by class*

@@

@@

>>>>

1. Illustration of problem

## Derivation of Multinomial Filter

@@

@@

## Class-based Ensemble adjustment

## VSEM Description

## Simulated data experiments

@@

Things to try:
* Run just the Assimilation step for simulated disturbance that vary in magnitude, process error, observation error, etc to detect sensitivity
* Approximation of not simulating disturbed ensemble members and relying SOLEY on ensemble adjustment
* Analytical approximation of sequential beta-binomial then KF vs full joint model

## Calibration of VSEM
*

[@BayesianTools]

## Comparison to real-world data

* LandTrendr tile
* Construction of frequency and transition matrix priors
